Finiteness in Cubical Type Theory

### 2020-01-20 paper.tex
more

  We also prove closure properties for each finite type, and classify them
  topos-theoretically.
  For instance we show that the category of decidable Kuratowski finite sets
  (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.

@@

  We formalise our work in Cubical Agda, and we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying),
  and demonstrate how it can be used to both prove properties and synthesise
  full functions given desired properties.
  \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
    Dependent Types \and Finiteness \and Topos \and Kuratowski finite}
\end{abstract}

@@

All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in the
final section, where we implement a library for proof search, based on
omniscience and exhaustibility.
While proof search based on finiteness is not new, implementing it here does
give us an opportunity to demonstrate some actual \emph{computation} in a
univalent setting.
Many of the finiteness predicates are built using univalence, so for the proof
search functionality to work at all we require computational content in the
univalence axiom.
Furthermore, the extensionality afforded to us by HoTT means that we can have
\(\Pi\) types in the domain of the search, meaning we can synthesise full
functions from specifications alone.
\subsection{Related Work}
\subsection{Notation}
\todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
\todo[inline]{Paths, Univalence, isomorphisms}
\todo[inline]{HITs}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (Kuratowski finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Kuratowski Finiteness}
The first thing we must define is a representation of subsets.
\begin{rm-definition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{aligned}
      \mathcal{K}(A) =&
      \; [] : \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{aligned}
  \end{equation}
\end{rm-definition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

@@

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{rm-definition}[Membership of \(\mathcal{K}\)]
  First, we need to provide equations for the two point constructors.
  \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
  + (x \in \mathit{xs}) \rVert \).
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(+\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{rm-definition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{rm-definition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{rm-definition}
\todo[inline]{Kf is a proposition, circle is Kf}

\subsection{Split Enumerable}
\begin{rm-definition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\end{rm-definition}
In this, and the rest of the definitions in the paper, we tend to prefer
list-based notions of finiteness.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{rm-lemma}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{rm-lemma}
We have proven this statement in our formalisation, but it is worth pointing out
that the proof is nearly trivial: after inlining the definitions of
\(\mathbf{List}\), \(\in\), and \(\twoheadrightarrow!\), it becomes clear that
the two terms are basically the same anyway, modulo the reassociation of
\(\Sigma\).

Split enumerability is quite a strong predicate, implying both decidable
equality and a total order on the underlying type.
This

\subsection{Manifest Bishop}
\subsection{Manifest Enumerable}
\subsection{Cardinal}
\section{Relations Between Each Finiteness Definition}
\section{Topos}
\begin{itemize}
  \item Complete (pullbacks + terminal object)
\end{itemize}
\section{Countably Infinite Types}
\section{Search}

@@

% Before giving the definition of the predicate, we will first define lists (and
% membership thereof): we have chosen a \emph{container}-based definition for this
% work.
% \begin{rm-definition}[Container] \label{container-def}
%   A container \cite{abbottContainersConstructingStrictly2005} is a pair
%   \(S \triangleright P\) where \(S\) is a type, the elements of which are called
%   the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
%   the elements of \(P(s)\) are called the \emph{positions} of a container.
%   We ``interpret'' a container into a functor defined like so:
%   \begin{equation}
%     \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
%   \end{equation}
%   Membership of a container can be defined like so:
%   \begin{equation}
%     x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
%   \end{equation}
%   Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
% \end{rm-definition}
% \begin{rm-definition}[Lists]
%   \begin{equation}%
%     \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
%   \end{equation}
% \end{rm-definition}
% \begin{rm-definition}[\mathbf{Fin}]
%   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
%   We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
%   \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
% \end{rm-definition}

@@

% Finally, we can define formally split enumerability.
% \subsection{Split Surjections}
% \begin{rm-definition}[Surjections]
%   \begin{multicols}{2}
%     \begin{equation}
%       A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
%     \end{equation} \break
%     \begin{equation}
%       A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
%     \end{equation}
%   \end{multicols}
%   Where the definitions of surjective are taken from \cite[definition
%   4.6.1]{hottbook}.
% \end{rm-definition}
% \begin{rm-theorem} \label{split-enum-surj}
%   Split enumerability is equivalent to a split surjection from a finite prefix
%   of the natural numbers.
%   \begin{equation}
%     \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
%   \end{equation}
% \end{rm-theorem}
% \begin{proof}
%   After sufficient inlining, it emerges that our goal is simply a reassociation.
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
%   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
%   have written out the extra lines for clarity alone.
% \end{proof}
% \begin{rm-lemma} \label{split-enum-discrete}
%   Any split enumerable type has decidable equality (is discrete).
% \end{rm-lemma}
% \begin{proof}
%   We use a corollary that if there is a split-surjection from \(A\) to \(B\),
%   and \(A\) is discrete, then \(B\) is also discrete.
% \end{proof}
% By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
% this implies split enumerable types must all be sets.
% \subsection{Closure}
% In this section we will prove closure under various operations for split
% enumerable sets.
% We are working towards a topos proof, which requires us to prove closure under
% a variety of operations: for now, we only have enough machinery to demonstrate
% the semiring operations, and dependent sums.
% In order to show closure under exponentials (function arrows), we will need an
% equivalence with \(\mathbf{Fin}\), which will be provided in
% section~\ref{manifest-bishop}.
% \begin{rm-lemma} \label{split-enum-bool-top-bot}
%   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
% \end{rm-lemma}
% \begin{proof}
%   Each of these types clearly has a finite number of elements (2, 1, and 0,
%   respectively), and furthermore has a straightforward enumeration.
% \end{proof}
% \begin{rm-theorem} \label{split-enum-sigma}
%   Split-enumerability is closed under \(\Sigma\).
%   \begin{equation} \label{split-enum-sigma-clos-sig}
%     \frac{
%       \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
%     }{
%       \mathcal{E}!(\Sigma (x : A) , U(x))
%     }
%   \end{equation}
% \end{rm-theorem}
% \begin{proof}
%   To obtain the support list, we concatenate the support lists of all the proofs
%   of split-finiteness for \(U\) over the support list of \(E_A\).
%   In Agda:
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
%   ``do-notation'' is available to us as we're working in the list monad.
% \end{proof}
% Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
% as these types can be defined in terms of \(\Sigma\).
% \begin{align}
%   A \times B &= \Sigma {(x : A)} , B \\
%   A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
% \end{align}
% \section{Manifest Bishop Finiteness} \label{manifest-bishop}
% Where split enumerability was the enumeration form of a split surjection from
% \(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
% \emph{equivalence} with \(\mathbf{Fin}\).

>>>>

  We also prove closure properties for each finite type, and classify them
  topos-theoretically.
  This includes a proof that the category of decidable Kuratowski finite sets
  (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.

@@

  We formalise our work in Cubical Agda, where we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying).
  Through this library we demonstrate a number of uses for the computational
  content of the univalence axiom, including searching for and synthesising
  functions.
  \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
    Dependent Types \and Finiteness \and Topos \and Kuratowski finite}
\end{abstract}

@@

All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in the
final section, where we implement a library for proof search, based on
omniscience and exhaustibility.
While proof search based on finiteness is not new, implementing it here does
give us an opportunity to demonstrate some actual \emph{computation} in a
univalent setting.
Many of the finiteness predicates are built using univalence, so for the proof
search functionality to work at all we require computational content in the
univalence axiom.
Furthermore, the extensionality afforded to us by HoTT means that we can have
\(\Pi\) types in the domain of the search, meaning we can synthesise full
functions from specifications alone.
\subsection{Related Work}
\subsection{Notation}
\todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
\todo[inline]{Paths, Univalence, isomorphisms}
\todo[inline]{HITs}
\todo[inline]{Surjections}
\todo[inline]{Containers}
\todo[inline]{Fin}
\todo[inline]{Biconditional, Isomorphism, Equivalence, Equals}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Kuratowski Finiteness}
Much work has already been done on Kuratowski finiteness in HoTT in
\cite{fruminFiniteSetsHomotopy2018}.
As a result, we will not needlessly repeat proofs, rather we will just give a
brief introduction to the topic and point out where our treatment differs.

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{aligned}
      \mathcal{K}(A) =&
      \; [] : \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{aligned}
  \end{equation}
\end{romdefinition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

@@

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{romdefinition}[Membership of \(\mathcal{K}\)]
  Membership is defined by pattern-matching on \(\mathcal{K}\).
  The two point constructors are handled like so:
  \begin{equation}
    \begin{alignedat}{2}
      x \in&& \; []                      &= \bot \\
      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
    \end{alignedat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(\uplus\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{romdefinition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{romdefinition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{romdefinition}

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\subsection{Split Enumerable}
\begin{romdefinition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \equiv \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\todo[inline]{Put refs for each ``definition of''}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    & \text{Definition of } \mathcal{E}!
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    & \text{Definition of } \in
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of sp-surj}
    \\
    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}

In our formalisation, the proof is a single line: most of the steps above are
simple expansion of definitions.
The only step which isn't definitional equality is the reassociation of
\(\Sigma\).

Split enumerability implies decidable equality on the underlying type.
To prove this, we will make use of the following lemma, proven in the
formalisation:
\begin{romlemma} \label{discrete-surj}
  \begin{equation}
    \frac{
        A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
      }{
       \text{Discrete}(B) 
      }
  \end{equation}
\end{romlemma}
\begin{romlemma}
  Every split enumerable type is discrete.
\end{romlemma}
\begin{proof}
  Let \(A\) be a split enumerable type.
  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
  \(\mathbf{Fin}(n)\) for some \(n\).
  Also, by lemma \todo[noline]{Fin is discrete}, we know that
  \(\mathbf{Fin}(n)\) is discrete.
  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
  \qed
\end{proof}
\subsection{Manifest Bishop}
\begin{romdefinition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
  \end{equation}
\end{romdefinition}
The only difference between manifest Bishop finiteness and split enumerability
is the membership term: here we require unique membership (\(\inunique\)),
rather than simple membership (\(\in\)).

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    & \text{Definition of } \mathcal{B}
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    & \text{Definition of } \inunique
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{\cite[theorem 4.4.5]{hottbook}}
    \\
    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}
\subsection{Manifest Enumerable}
\begin{romdefinition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{romdefinition}
Again, the only difference with this type and split enumerability is the
membership proof.
Here we have propositionally truncated it.

% We will use the circle as an example non-set type which is nonetheless
% manifestly enumerable.
% \begin{romtheorem}
%   The circle \(S^1\) is manifestly enumerable.
% \end{romtheorem}
% \begin{proof}
%   As the cover proof is a truncated proposition, we need only consider the point
%   constructors, making this poof the same as the proof of split enumerability on
%   \(\top\).
% \end{proof}
% \subsection{Surjections}
% This predicates relation to surjectivity is much the same as split
% enumerability's relation to \emph{split} surjectivity.
% \begin{romlemma}
%   A proof of manifest enumerability is equivalent to a surjection from a finite
%   prefix of the natural numbers.
% \end{romlemma}
\subsection{Cardinal}
\section{Relations Between Each Finiteness Definition}
\section{Topos}
\begin{itemize}
  \item Complete (pullbacks + terminal object)
\end{itemize}
\section{Countably Infinite Types}
\section{Search}

@@

% Before giving the definition of the predicate, we will first define lists (and
% membership thereof): we have chosen a \emph{container}-based definition for this
% work.
% \begin{romdefinition}[Container] \label{container-def}
%   A container \cite{abbottContainersConstructingStrictly2005} is a pair
%   \(S \triangleright P\) where \(S\) is a type, the elements of which are called
%   the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
%   the elements of \(P(s)\) are called the \emph{positions} of a container.
%   We ``interpret'' a container into a functor defined like so:
%   \begin{equation}
%     \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
%   \end{equation}
%   Membership of a container can be defined like so:
%   \begin{equation}
%     x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
%   \end{equation}
%   Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
% \end{romdefinition}
% \begin{romdefinition}[Lists]
%   \begin{equation}%
%     \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
%   \end{equation}
% \end{romdefinition}
% \begin{romdefinition}[\mathbf{Fin}]
%   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
%   We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
%   \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
% \end{romdefinition}

@@

% Finally, we can define formally split enumerability.
% \subsection{Split Surjections}
% \begin{romdefinition}[Surjections]
%   \begin{multicols}{2}
%     \begin{equation}
%       A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
%     \end{equation} \break
%     \begin{equation}
%       A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
%     \end{equation}
%   \end{multicols}
%   Where the definitions of surjective are taken from \cite[definition
%   4.6.1]{hottbook}.
% \end{romdefinition}
% \begin{romtheorem} \label{split-enum-surj}
%   Split enumerability is equivalent to a split surjection from a finite prefix
%   of the natural numbers.
%   \begin{equation}
%     \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   After sufficient inlining, it emerges that our goal is simply a reassociation.
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
%   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
%   have written out the extra lines for clarity alone.
% \end{proof}
% \begin{romlemma} \label{split-enum-discrete}
%   Any split enumerable type has decidable equality (is discrete).
% \end{romlemma}
% \begin{proof}
%   We use a corollary that if there is a split-surjection from \(A\) to \(B\),
%   and \(A\) is discrete, then \(B\) is also discrete.
% \end{proof}
% By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
% this implies split enumerable types must all be sets.
% \subsection{Closure}
% In this section we will prove closure under various operations for split
% enumerable sets.
% We are working towards a topos proof, which requires us to prove closure under
% a variety of operations: for now, we only have enough machinery to demonstrate
% the semiring operations, and dependent sums.
% In order to show closure under exponentials (function arrows), we will need an
% equivalence with \(\mathbf{Fin}\), which will be provided in
% section~\ref{manifest-bishop}.
% \begin{romlemma} \label{split-enum-bool-top-bot}
%   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
% \end{romlemma}
% \begin{proof}
%   Each of these types clearly has a finite number of elements (2, 1, and 0,
%   respectively), and furthermore has a straightforward enumeration.
% \end{proof}
% \begin{romtheorem} \label{split-enum-sigma}
%   Split-enumerability is closed under \(\Sigma\).
%   \begin{equation} \label{split-enum-sigma-clos-sig}
%     \frac{
%       \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
%     }{
%       \mathcal{E}!(\Sigma (x : A) , U(x))
%     }
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   To obtain the support list, we concatenate the support lists of all the proofs
%   of split-finiteness for \(U\) over the support list of \(E_A\).
%   In Agda:
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
%   ``do-notation'' is available to us as we're working in the list monad.
% \end{proof}
% Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
% as these types can be defined in terms of \(\Sigma\).
% \begin{align}
%   A \times B &= \Sigma {(x : A)} , B \\
%   A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
% \end{align}
% \section{Manifest Bishop Finiteness} \label{manifest-bishop}
% Where split enumerability was the enumeration form of a split surjection from
% \(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
% \emph{equivalence} with \(\mathbf{Fin}\).

% \begin{romdefinition}[Manifest Bishop Finiteness]
%   \begin{equation}
%     \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
%   \end{equation}
% \end{romdefinition}
% The only difference between this predicate and split enumerability is the list
% membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
% to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
% \begin{romdefinition}[Unique Membership]
%   We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
%   membership in that list is a \emph{contraction} \cite[definition
%   3.11.1]{hottbook}; i.e. its membership proof exists, and all such proofs are
%   equal.
%   \begin{equation}
%     x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
%   \end{equation}
% \end{romdefinition}
% By the definition of a contraction, we can always recover the underlying
% membership proof, meaning that we can always derive split enumerability from
% manifest Bishop finiteness.
% \subsection{Equivalence}
% \begin{romlemma} \label{bishop-equiv}
%   A proof of manifest Bishop finiteness is equivalent to an equivalence with a
%   finite prefix of the natural numbers.
%   \begin{equation}
%     \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
%   \end{equation}
% \end{romlemma}
% \begin{proof}
%   There are many equivalent definitions of equivalence in HoTT.
%   Here we take the version preferred in the Cubical Agda library: contractible
%   maps \cite[definition 4.4.1]{hottbook}.
%   Because of the parallels between contractible maps and split surjections,
%   the proof proceeds much the same as \ref{split-enum-surj}.
%   In other words, the definition of Bishop finiteness is itself a reassociation
%   of a contractible map.
% \end{proof}
% \subsection{Relationship to Split Enumerability}
% \begin{romtheorem} \label{split-enum-to-manifest-bishop}
%   Any split enumerable set is manifest Bishop finite.
% \end{romtheorem}
% \begin{proof}
%   From proposition~\ref{split-enum-discrete} we can derive decidable equality on
%   \(A\), and using this we can define a function (called \(\text{uniques}\)) which filters
%   out duplicates from lists of \(A\)s.
%   This gives us our support list.
%   To generate the cover proof it suffices now to prove the following:
%   \begin{equation}
%     \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
%   \end{equation}
% \end{proof}
% \subsection{Closure}
% Proving equal strength of split enumerability and manifest Bishop finiteness
% allows us to carry all of the previous proofs of closure over to manifest Bishop
% finite sets (and vice-versa).
% Missing from our previous proofs was a proof of closure of functions.
% We remedy that here.
% \begin{romtheorem}
%   Manifest bishop finiteness is closed over dependent functions
%   (\(\prod\)-types).
%   \begin{equation}
%     \frac{
%       \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
%     }{
%       \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
%     }
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   This proof is essentially the composition of two transport operations, made
%   available to us via univalence.
